---
title: "Tidybiology: ggplot2 (Answers)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Load packages

# install.packages("tidybiology")
library(tidyverse)
# install.packages("ggExtra")
library(ggExtra)
library(tidybiology)

# Load data
data(happy_full)
data(happy_select)
```

## Visualizing data from the World Happiness Report

Now we move to the fun part - making pretty plots!

## Plotting distributions

Let's begin by getting a sense of the overall distribution of `ladder_score` in the `happy_select` data frame

```{r}
happy_select %>% 
  ggplot(aes(ladder_score)) +
  geom_histogram()
```

### Making improvements

This plot is fine, but it's a little chunky. Let's try a different geom - `geom_density()` - to see what we get

```{r}
happy_select %>% 
  ggplot(aes(ladder_score)) +
  geom_density()
```

### Density by region

What does the `ladder_score` distribution look like for each region? There are many ways to visualize this. Let's first plot all the distributions on one plot  

```{r}
happy_select %>% 
  ggplot(aes(ladder_score, fill = region)) +
  geom_density()
```

Nice! One small problem with this plot is that distributions are overlapping, making it difficult to visualize everything. Which argument can you adjust to improve this plot? Does it go within `aes()` or outside? Why?  

### Faceting

Let's consider an alternative way to plot the distributions. Go ahead and use the `facet_wrap()` function to do this

```{r}
happy_select %>% 
  ggplot(aes(ladder_score)) +
  geom_density() +
  facet_wrap(~region)
```

## Scatterplots

### Are rich countries "happier"?

We can use scatterplots to visualize the relationship between two variables. For example, let's take a look at the relationship between `ladder_score` and `logged_gdp_per_capita` (this relationship might be obvious, but maybe we'll be surprised!) 

```{r}
happy_select %>% 
  ggplot(aes(logged_gdp_per_capita, ladder_score)) +
  geom_point()
```

### Emphasizing relationships

It's clear from the previous scatterplot that relationship between `ladder_score` and `logged_gdp_per_capita` is positive, and it looks pretty linear. To highlight this, we can use the `geom_smooth()` geom (remember to set the `se` argument to equal `FALSE` to just get a straight line) 

```{r}
happy_select %>% 
  ggplot(aes(ladder_score, logged_gdp_per_capita)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

## Making things prettier

The plots we've made, while nice, look a little...basic. Let's try to make some improvements. We'll use the last plot  

Add a title, subtitle, x-axis label, and y-axis label. Also change the overall look of the plot using your favourite `theme_*()` function  

```{r}
happy_select %>% 
  ggplot(aes(ladder_score, logged_gdp_per_capita)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = "Does Money == Happiness?",
       subtitle = "Plotting Ladder Score against Log GDP Per Capita",
       x = "Ladder Score",
       y = "Log GDP Per Capita") +
  theme_minimal()
```

## Adding marginal plots

Add histograms for `logged_gdp_per_capita` and `ladder_score` to the previous plot (to the top and to the right, respectively) 

**Hint:** Use the `ggMarginal()` function from the `ggExtra` package. To learn more about this function, run `?ggMarginal`
```{r}
p1 <- happy_select %>% 
        ggplot(aes(ladder_score, logged_gdp_per_capita)) +
        geom_point() +
        geom_smooth(se = FALSE) +
        labs(title = "Does Money == Happiness?",
             subtitle = "Plotting Ladder Score against Log GDP Per Capita",
             x = "Ladder Score",
             y = "Log GDP Per Capita") +
        theme_minimal()

ggMarginal(
  p = p1,
  type = 'histogram'
)
```

## Applying the "final" touch

Make the x- and y-axis text bold

```{r}
p1 <- happy_select %>% 
        ggplot(aes(ladder_score, logged_gdp_per_capita)) +
        geom_point() +
        geom_smooth(se = FALSE) +
        labs(title = "Does Money == Happiness?",
             subtitle = "Plotting Ladder Score against Log GDP Per Capita",
             x = "Ladder Score",
             y = "Log GDP Per Capita") +
        theme_minimal() +
        theme(axis.text = element_text(face = "bold"))

ggMarginal(
  p = p1,
  type = 'histogram'
)
```

Save the last plot
```{r}
ggsave(file = "my_plot.png", ggMarginal(p = p1, type = 'histogram'))
```

## For more help

Run the following to access the Ggplot2 vignette

```{r}
browseVignettes("ggplot2")
```

