---
title: "Tidybiology: Class 5 Report Exercise" 
subtitle: "My favourite country is `r params$country_name`"
output: html_document  
params: 
  country_name: "India"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Load packages

# install.packages("tidybiology")
# remotes::install_github("hadley/emo")
library(tidyverse)
library(tidybiology)
library(emo)

# Load data
data(happy_full)
```

```{r include = FALSE}
# Obtain information for your country
my_country <- filter(happy_full, country_name == params$country_name)
```

`r my_country$country_name` `r emo::ji(str_to_lower(params$country_name))` is a country in `r my_country$region`. Residents of `r my_country$country_name` have a life expectancy of `r round(my_country$healthy_life_expectancy)`, which is `r abs(my_country$healthy_life_expectancy - mean(happy_full$healthy_life_expectancy, na.rm = TRUE) %>% round())` years **`r if_else((my_country$healthy_life_expectancy - mean(happy_full$healthy_life_expectancy, na.rm = TRUE)) > 0, "greater", "less")`** than the average global life expectancy of `r round(mean(happy_full$healthy_life_expectancy))`.

```{r provenance, include=FALSE}
#Session information for provenance and reproducibility
session_provenance()
```
