---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidybiology

<!-- badges: start -->
<!-- badges: end -->

This package accompanies a workshop-style class that provides an introduction to the emerging field of Data Science in R, including data analysis and visualization, with a particular focus on its utility for biological insight. In class, students will be provided with biological datasets using this package, and introduced to the `tidyverse` and `code` used to examine data. In the first half of each class, students will be lectured on methods and shown demonstrations; in the second half of each class, students will use tools to analyze real data. Methods for filtering, sorting, and transforming data will be discussed along with visualization tools and options. Particular attention will be paid to code interpretation and data provenance methods by learning to generate reproducible data output files. For a final project, students will be given a new dataset to analyze using the tools learned during the course, and will share findings with the class in a short oral presentation. Although specific datasets will be used for analysis in class, this workshop will provide broadly applicable tools to reproducibly analyze and visualize data across the biological sciences.  

## Installation

You cannot yet install the released version of tidybiology from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("tidybiology")
```

So in the meantime, use the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hirscheylab/tidybiology")
```
## Example

To see the datasets currently contained within this package:

```{r example}
#to load the library
library(tidybiology)

#to see the datasets within this package
data(package = "tidybiology")

#to load a dataset
data(simplechromosome)

#simple plot
plot(simplechromosome, xlab = "Chromosome", ylab = "Number of Basepairs")
```
