---
title: 'Tidybiology: dplyr homework'
author: "Devang Thakkar"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
```

## Workshop

You will be using the `scRNAseq` dataset to answer all the following questions

Add your answers (text and/or code chunks as required) in the space after each question

The `scRNAseq` dataset can be found in the scRNAseq package. To load it, run the following

```{r}
devtools::install_github("devangthakkar/scRNAseq")
library(scRNAseq)
```

## Please answer the following questions

What type of object is `cell_1`?

```{r, eval=TRUE, echo = TRUE}

```

Filter the dataset to only include genes that are specifically expressed in the brain or in the heart and assign it to a new dataset `organ_data`. How many genes does `organ_data` have?

```{r}

```

Using the `organ_data` dataset, select the following columns - `gene_name`, `organ`, and any column with the word `cell` in the column name. Store the result into a new dataset `filtered_data`. How many columns does `filtered_data` have?

```{r}

```

Sort the `filtered_data` dataset from Z to A based on `gene_name`. Store the result into a new dataset `sorted_data`. 

```{r}

```

Using the `sorted_data` dataset, we now want to look at what organ does `cell_1` originate from. Group the data set by the `organ`, and summarize the mean counts for `cell_1`

```{r}

```

BONUS: Identify the origin to which all the 10 cells belong. You will need to use the `across` functionality to select all columns that contain the word `cell`. Check out the documentation here: https://dplyr.tidyverse.org/reference/across.html

```{r}
# filtered_data %>% group_by(organ) %>% summarise(across(contains("cell"), mean))
```
