---
title: "ggplot2 HW"
author: "Junqi Lu"

output:
  html_document: 
    
    toc: true 
    toc_float: 
      collapsed: false 
      smooth_scroll: false 
      
    number_sections: true 
    
    theme: paper 
    
    output: 
      fig_width: 10 
      fig_height: 8 
      fig_caption: true 
      dev: svg 
    
    #Dataframe printing
    df_print: paged 
    
    code_folding: hide 
  
  pdf_document: default
---

# Create a style you like to use later for all the other plots
You need to load __ggplot2__ library before you define your style in a list. 
```{r}

```


# Load data
Load scRNAseq data from our library below. 
```{r}

```

# Preview data structure by the method you prefer 
```{r}

```

# Generate plots
* For the questions below, feel free to use the __ggplot2 builder__ as an assistance.
* Remember to label you axes and title your legends and plots properly.


## Which gene type has a bigger transcript length? 
Choose the range of transcript length that you think is reasonable to be used for the plot.  

* The length might not make sense and thus can potentially be an artifact that you want to remove.    
* You may need to adjust the plot width or turn the axis texts on x-axis with an angle to avoid overlapping.

```{r}

```

## Which organ (excluding the "other" category) has a higher average transcript length? 
Excluding extreme outliers by strategically filtering the data can help you focus more. 
Also use __ggsignif__ to label out the p values of 2 organs that you're interested in color __red__ so people can easily observe whether the difference is significant. 
```{r}

```

## For the gene type that is expressed in more organ types, in which organ does that gene type have the highest average GC content? 
Please use the __viridis__ coloring scheme here. 
```{r}

```

## Pick 2 cells that you like to compare
Let's focus on the organ types other than "other". Do you think these 2 cells are more likely to be of the same cell type or different cell types? 

Add marginal histograms around both axes by __ggExtra__. Adding a y=x line on the plot can help you approximate the correlation of these 2 cells' RNA. 

* If same cell type, which organ(s) are they most likely coming from?
* If different cell types, which organ(s) does each of the 2 cells might come from? 
```{r}

```
 

## Explore ggplot 2 extensions
Discover a new ggplot2 extension from https://exts.ggplot2.tidyverse.org/gallery/ and try to use it below in a plot that you already made previously. 

Suggestion: use __gghalves__ extension to change the right half of geom_boxplots into half violin plots for information about distribution. 
```{r}

```

