---
title: "Data Transformation with dplyr & stringr + ggplot2 HW"
author: "Junqi Lu"

output:
  html_document: 
    
    toc: true 
    toc_float: 
      collapsed: false 
      smooth_scroll: false 
      
    number_sections: true 
    
    theme: paper 
    
    output: 
      fig_width: 10 
      fig_height: 8 
      fig_caption: true 
      dev: svg 
    
    #Dataframe printing
    df_print: paged 
    
    code_folding: show 
  
  pdf_document: default
---
# Load data
```{r}
library('scRNAseq')
scRNAData <- scRNAseq #Because ggplot2 builder needs the dataset to be in the Environment
```

## Preview data structure 
```{r}
library('dplyr')
glimpse(scRNAData)
```

```{r}
head(scRNAData,4) #Dummy datasets will be generated from these 4 rows
```

# Generate dummy df
```{r}
heart <- scRNAData[scRNAData$organ=='heart',]
brain <- scRNAData[scRNAData$organ=='brain',]
kidney <- scRNAData[scRNAData$organ=='kidney',]
```

# Does the original df have any transcript appeared in more than 1 organ?

* Clue for more than 2 df join(): try to use %>%
```{r}

```


# Assembly the df from 3 organs together in a proper way and name the new df as three_organ_df
```{r}

```

# FIlter the genes from heart out from three_organ_df 

* You can use either a join function or a negated %in% (__`%!in%` <- Negate(`%in%`)__)
* If you want to use a join function, make sure you understand which is the "left" and which is the "right"
```{r}

```

# Conditional coloring in ggplot by stringr pattern detect 
## What gene types do you have in scRNAData? 
```{r}
sort(table(scRNAData$gene_type), decreasing = TRUE)
```

## What is the average transcript lengthn for non-pseudogenes from brain? Make a plot to answer this. 

Additional requirements: 

* Filter the scRNAData by transcript length by a range that you think is reasonable to remove potential artifacts
* Change the x-axis label of a boxplot to __Pseudogenes__ if its gene type belongs to pseudogenes or __Non-pseudogenes__ if not
* Use the organ label to fill in the color of boxplots. If a gene is from brain, label it as __Brain-related__; if not from brain, label it as __Other-organs-related__

```{r}

```



