---
title: "Introduction to tidybiology"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
library(tidybiology)
datasets <- data(package = "tidybiology")$result[, "Item"]
```
## `tidybiology`: An Introduction to Biological Data Science in R
Tidybiology is workshop-style series of lectures that provides an introduction to the emerging field of Data Science in R, including data analysis and visualization, with a particular focus on its utility for biological insight. This series is designed to be either self-taught or lectures for in-person learning. The `tidybiology` package provides biological datasets to be used along side the `tidyverse` R packages and code.  
  
### Viewing datasets
This package currently contains `r length(datasets)` datasets. To see the datasets currently contained within the `tidybiology` package:  
```{r packages, echo=TRUE}
library(tidybiology) #load the library
data(package = "tidybiology")
```

```{r printpackages, echo=FALSE}
print(datasets)
```
  
### Using datasets
To load a dataset, simple use the `data()` function:
```{r load}
data(simplechromosome)
```
  
Make a simple plot, by calling the loaded dataframe:  
```{r plot}
plot(simplechromosome, xlab = "Chromosome", ylab = "Number of Basepairs")
```
