---
title: "intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
library(tidybiology)
datasets <- data(package = "tidybiology")$result[, "Item"]
```
## TidyBiology: An Introduction to Biological Data Science in R
Tidybiology is workshop-style series that provides an introduction to the emerging field of Data Science in R, including data analysis and visualization, with a particular focus on its utility for biological insight. Students will be provided with biological datasets using the `tidybiology` package, and introduced to the `tidyverse` R packages and code to examine data.  
  
### Viewing datasets
This package currently contains ```r length(datasets)``. To see the datasets currently contained within the `tidybiology` package:  
```{r packages, echo=TRUE}
library(tidybiology) #load the library
data(package = "tidybiology")
```

```{r printpackages, echo=FALSE}
print(datasets)
```
  
### Using datasets
To load a dataset, simple use the `data()` function:
```{r load}
data(simplechromosome)
```
  
Make a simple plot, as you would:  
```{r plot}
plot(simplechromosome, xlab = "Chromosome", ylab = "Number of Basepairs")
```
