---
title: 'Tidybiology: dplyr homework'
author: "Devang Thakkar"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
hook_output <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  lines <- options$output.lines
  if (is.null(lines)) {
    return(hook_output(x, options))  # pass to default hook
  }
  x <- unlist(strsplit(x, "\n"))
  more <- "..."
  if (length(lines)==1) {        # first n lines
    if (length(x) > lines) {
      # truncate the output, but add ....
      x <- c(head(x, lines), more)
    }
  } else {
    x <- c(more, x[lines], more)
  }
  # paste these lines together
  x <- paste(c(x, ""), collapse = "\n")
  hook_output(x, options)
})
```

## Workshop

You will be using the `scRNAseq` dataset to answer all the following questions

Add your answers (text and/or code chunks as required) in the space after each question

The `scRNAseq` dataset can be found in the scRNAseq package. To load it, run the following

```{r}
devtools::install_github("devangthakkar/scRNAseq")
library(scRNAseq)
```

## Please answer the following questions

What type of object is `cell_1`?

```{r, eval=TRUE, echo = TRUE}
class(scRNAseq$cell_1)
```

Filter the dataset to only include genes that are specifically expressed in the brain or in the heart and assign it to a new dataset `organ_data`. How many genes does `organ_data` have?

```{r}
organ_data <- scRNAseq %>%
              filter(organ == "heart" | organ == "brain")
organ_data
```

Using the `organ_data` dataset, select the following columns - `gene_name`, `organ`, and any column with the word `cell` in the column name. Store the result into a new dataset `filtered_data`. How many columns does `filtered_data` have?

```{r}
filtered_data <- organ_data %>%
                 select(gene_name | organ | contains("cell"))
filtered_data
```

Sort the `filtered_data` dataset from Z to A based on `gene_name`. Store the result into a new dataset `sorted_data`. 

```{r}
sorted_data <- filtered_data %>% 
               arrange(desc(gene_name))
sorted_data
```

Using the `sorted_data` dataset, we now want to look at what organ does `cell_1` originate from. Group the data set by the `organ`, and summarize the mean counts for `cell_1`

```{r}
sorted_data %>% group_by(organ) %>%
                summarize(mean(cell_1))
```

BONUS: Identify the origin to which all the 10 cells belong. You will need to use the `across` functionality to select all columns that contain the word `cell`. Check out the documentation here: https://dplyr.tidyverse.org/reference/across.html

```{r}
sorted_data %>% group_by(organ) %>% 
                summarise(across(contains("cell"), mean))
```