---
title: "Tidybiology: dplyr"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Load packages

# install.packages("tidybiology")
library(tidyverse)
# install.packages("fontawesome")
library(fontawesome)
library(tidybiology)

# Load data
data(happy_full)
data(happy_select)
```

## Analysing the World Happiness Report `r fa("globe-europe")`

In this exercise, you will apply what you've learned in class to perform exploratory data analysis (EDA) on the **World Happiness Report**    

This dataset was downloaded from the website [Kaggle](https://www.kaggle.com/ajaypalsinghlo/world-happiness-report-2021?select=world-happiness-report.csv). We will use the **2021** data in this exercise. This dataset is stored in an object called `happy_full` 

###  

In this exercise, you will practice:

* gaining quick insight into the type of data this dataset contains
* using functions from the Dplyr package to wrangle your data and obtain useful summaries

## Take a look at your data

### What does the dataset look like?

A couple useful things to know about your dataset are -  
- The number of rows and columns
- The types of variables the dataset contains  

What function can you use to get this information?

```{r}
dim(happy_full)
glimpse(happy_full)
```

### Double data type

We can see that `happy_full` contains many variables that are of type double. 

The `double` data type refers to which of the following?  

- A string
- An integer
- A number with a decimal point
- A factor

## Selecting Variables 

The `happy_full` dataset contains many variables. This gives us the chance to practice our `select()`-ing skills!

### Simple selects

Let's warm up by performing some basic select operations  

How would you select just the columns `region` and `ladder_score`?  

```{r}
happy_full %>% 
  select(region, ladder_score)
```

Now select everything between (and including) `social_support` and `generosity`

```{r}
happy_full %>% 
  select(social_support:generosity)
```

### Slightly-more-difficult selects  

Let's try something more challenging now. Select all variables that **do not** have underscores in their names

**Hint:** You'll need a helper function. Also, don't forget `!`

```{r}
happy_full %>% 
  select(!contains("_"))
```

Helper functions can be really...helpful! Ok, no we're ready to select the variables we will need for the rest of this exercise. Create a new dataframe called `happy_df` that contains the following variables (in the specified order!) - `country_name`, `region`, `ladder_score_in_dystopia`, `logged_gdp_per_capita`, `social_support`, `healthy_life_expectancy`, `freedom_to_make_life_choices`, `generosity`, `perceptions_of_corruption`  

Avoid simply typing out the names of all these variables. Add `glimpse(happy_df)` as the last line to see if you got the right answer  

```{r}

```

**Note:** `happy_df` hasn't actually been saved anywhere. So we will be using an identical dataset called `happy_select` for the rest of this exercise. 

## Filtering

`happy_select` contains both numeric and character variables, with lots and lots of observations (rows). This gives us a great opportunity to practice our filtering skills!  

### Simple filters

Say we're only interested in looking at data for countries in East Asia. How would we do this?

**Hint:** To see which major regions are in this dataset, run `unique(happy_select$region)` 

```{r}
happy_select %>% 
  filter(region == "East Asia")
```

### Filter%in%g

Now use `filter()` to only keep data for the following countries - Algeria, Belgium, India, Tunisia, and Uganda. Try to do this without writing multiple filter statements  

**Hint:** Try the `%in%` operator 

```{r}
happy_select %>% 
  filter(country_name %in% c("Algeria", "Belgium", "India", "Tunisia", "Uganda"))
```

### Numeric filtering

Finally, let's filter out information for countries that have a **below average** ladder score 

**Hint:** Use the base R `mean()` function 

```{r}
happy_select %>% 
  filter(ladder_score > mean(ladder_score))
```

## Creating new variables

### Creating normalized ladder scores

Create a new variable called `normalized_ladder_score` that contains the ladder scores for each county in `happy_full` divided by the ladder score in dystopia  

As a bonus, re-order the resulting data frame in descending order of `normalized_ladder_score`

```{r}
happy_select %>% 
  mutate(normalized_ladder_score = ladder_score/ladder_score_in_dystopia) %>% 
  arrange(desc(normalized_ladder_score))
```

### Discarding variables

Do the same as for the previous question but only keep the following columns -
`ladder_score`, `ladder_score_in_dystopia`, `normalized_ladder_score`

```{r}
happy_select %>% 
  mutate(normalized_ladder_score = ladder_score/ladder_score_in_dystopia) %>% 
  arrange(desc(normalized_ladder_score)) %>% 
  select(contains("ladder"))

# or
happy_select %>% 
  mutate(normalized_ladder_score = ladder_score/ladder_score_in_dystopia,
         .keep = "used") %>% 
  arrange(desc(normalized_ladder_score))
```

## Summary statistics

### Which regions are the happiest? 

Let's now find out which are the happiest regions in the world. We'll do this by working out the average `ladder_score` of all the countries in each region 

**Hint:** Remember the best friends, `group_by()` and `summarise()` 

```{r}
happy_select %>% 
  group_by(region) %>% 
  summarise(mean(ladder_score))
```

### Maxed out

Now let's work out the maximum value for each numeric variable, for each region. Save the output of this code as an object named `output` 

**Hint:** `across()` is helpful here. To learn more about this function, run 
`?across`

```{r}
output <- happy_select %>% 
            group_by(region) %>% 
            summarise(across(where(is.numeric), max))
```

Save `output` as a new file (e.g. csv)

```{r csv save, eval = FALSE, echo = TRUE}
write_csv(output, "dplyr_output.csv")
```

## For more help

Run the following to access the Dplyr vignette

```{r Dplyr vignette, eval = FALSE, echo = TRUE}
browseVignettes("dplyr")
```

