---
title: "Tidybiology: joins and stringr (Answers)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Load packages

# install.packages("tidybiology")
library(tidyverse)
# install.packages("ggExtra")
library(ggExtra)
library(tidybiology)

# Load data
data(happy)
data(happy_join)
data(happy_full)
```

## Joins

Let's now practice combining datasets by using the Join family of functions from the Dplyr package

The two datasets we will be combining are called `happy` and `happy_join`. These are both tiny datasets that will make it easier to understand what the `*_join()` functions are doing  

Let's first familiarize ourselves with these two datasets - 

`happy`
```{r happy, echo = FALSE}
happy %>% 
  knitr::kable()
```

`happy_join`  
```{r happy_join, echo = FALSE}
happy_join %>% 
  knitr::kable()
```

### `left_join()`

A requirement for all joins is the presence of a variable that is **common** to both datasets being joined. In the case of `happy` and `happy_join` this variable is `country_name` 

Let's perform a `left_join()` on this variable and then examine the output 

```{r left-join}
happy %>% 
  left_join(happy_join, by = "country_name")
```

Compare this output to `happy`. Notice that we now have additional information on healthy life expectancy for all the countries in `happy` *except* for Spain. The reason for the `NA` for Spain in this category is because this information is missing from `happy_join`  

So you can see how we can use `left_join()` to add new variable(s) to our dataset 

### `right_join()`

Now do a right join, recalling that this is syntactically identical to a left join except you replace "left" with "right"  

```{r right-join}
happy %>% 
  right_join(happy_join, by = "country_name")
```

Carefully look at the output. How is this different from a left join? In this output we only retain countries found in `happy_join`. Along with the `healthy_life_expectancy` variable, we also have three additional variables - `ladder_score`, `gdp`, and `social_support` - obtained from the `happy` dataset. Again, countries that are non-overlapping between these two datasets have NAs for these additional variables   

So you can think of a right join as being the inverse of a left join  

### `inner_join()`

Again, run the code for an inner join and we'll then take a look at the output  

```{r inner-join}
happy %>% 
  inner_join(happy_join, by = "country_name")
```

Here we see that we produce a dataset that only contains countries that are common to both `happy` and `happy_join`. This dataset also contains all the variables from both original datasets, and has the nice feature of not containing any missing values (i.e. NAs)

This is a useful join to use if you want your output to be complete and not contain any missing data 

### `full_join()`

Let's complete this section with the most complete join - the full join 

```{r full-join}
happy %>% 
  full_join(happy_join, by = "country_name")
```

As the name indicates, this is the most complete join that produces a dataset that contains all the information from both `happy` and `happy_join`. Use this if you don't want to discard any data during your data wrangling 

## Stringr 

### Country name lengths

Work out the number of characters (including spaces) for each country name in the `happy_full` data frame and populate a new column called `name_length` 

Which country has the longest name (in terms of number of characters)?

```{r}
happy_full %>% 
  mutate(name_length = str_count(country_name))
```

### Focusing on Europe 

Filter the `happy_full` data frame so that it only contains rows that correspond to European countries

```{r}
happy_full %>% 
  filter(str_detect(region, "Europe"))
```

### Europe or America 

Filter the `happy_full` data frame so that it only contains rows that correspond to European or American countries

```{r}
happy_full %>% 
  filter(str_detect(region, "Europe|America"))
```

### Filling in gaps

Some entries in the `region` column contain spaces in their names. Replace these with underscores

```{r}
happy_full %>% 
  mutate(region = str_replace_all(region, " ", "_"))
```

## For more help

Run the following to access the Stringr vignette

```{r}
browseVignettes("stringr")
```

